---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Your Name"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
library(tidyverse)
cookies <- read_csv("choc_chip_cookie_ingredients.csv")
glimpse(cookies)
```

```{python}
import pandas as pd
cookies = pd.read_csv("choc_chip_cookie_ingredients.csv")
cookies.head()
```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. Which ingredients appear in the most cookie recipes?
2. Do recipes with higher ratings use more or less sugar or butter?
3. What is the typical quantity of chocolate chips per recipe?
4. Are there ingredients that rarely appear together in the same recipe?
5. Do recipes from AllRecipes differ in ingredient quantities compared to Epicurious or hand-collected recipes?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library(skimr)
library(dplyr)
library(skimr)
library(dplyr)
skim_with(
  numeric = sfl(mean, sd, min, max, hist),
  character = sfl(n_unique, n_missing)
)

skim(cookies)

```

```{python}
from skimpy import skim

skim(cookies[['Quantity', 'Rating']])

skim(cookies[['Ingredient', 'Unit']])

```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

**The data shows that ingredient and unit columns are complete and standardized, with many unique ingredients. The rating is missing for about half of the recipes. This means we can analyze ingredients and quantities reliably, but ratings are incomplete**


### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.
```{r}
library(dplyr)
ingredient_props <- cookies |>
  group_by(Ingredient) |>
  summarise(count = n_distinct(Recipe_Index), .groups = "drop") |>
  mutate(proportion = count / n_distinct(cookies$Recipe_Index)) |>
  arrange(desc(count)) |>
  slice_head(n = 20)

ingredient_props

rare_ingredients <- cookies |>
  group_by(Ingredient) |>
  summarise(count = n_distinct(Recipe_Index), .groups = "drop") |>
  filter(count < 20) |>
  pull(Ingredient) |>
  paste(collapse = ", ")

rare_ingredients
```

```{python}
import pandas as pd

ingredient_counts = (
    cookies.groupby('Ingredient')['Recipe_Index']
    .nunique()
    .reset_index(name='count')
)
ingredient_counts['proportion'] = ingredient_counts['count'] / cookies['Recipe_Index'].nunique()
top20_ingredients = ingredient_counts.sort_values(by='count', ascending=False).head(20)
print(top20_ingredients)

rare_ingredients = ingredient_counts[ingredient_counts['count'] < 20]['Ingredient'].tolist()
rare_ingredients_str = ", ".join(rare_ingredients)
print(rare_ingredients_str)

```

[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.

```{r}
rare_ingredients <- cookies |>
  group_by(Ingredient) |>
  summarise(count = n_distinct(Recipe_Index), .groups = "drop") |>
  filter(count < 20) |>
  pull(Ingredient) |>
  paste(collapse = ", ")

rare_ingredients
```

```{python}
import pandas as pd

ingredient_counts = cookies.groupby("Ingredient")["Recipe_Index"].nunique()

rare_ingredients = ingredient_counts[ingredient_counts < 20].index.tolist()

rare_ingredients_str = ", ".join(rare_ingredients)

print(rare_ingredients_str)

```

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.
```{r}
library(ggplot2)
top_ingredients <- cookies |>
  group_by(Ingredient) |>
  summarise(num_recipes = n_distinct(Recipe_Index), .groups = "drop") |>
  arrange(desc(num_recipes)) |>
  slice_head(n = 20)

ggplot(top_ingredients, aes(x = reorder(Ingredient, num_recipes), y = num_recipes)) +
  geom_col(fill = "chocolate") +
  coord_flip() +
  labs(title = "Top 20 Most Common Ingredients in Cookie Recipes",
       x = "Ingredient",
       y = "Number of Recipes")

```

[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

**From the plot, the top ingredients across recipes are egg, vanilla and flour indicating that most recipes share a similar core base. The less common ingredients show where recipes differ and allow for variation in flavor and texture. In the future, it could be interesting to investigate whether the presence of these less common ingredients affects recipe ratings.**